---
title: "Cleaning Daily Checkin Data"
author: "Me"
date: "3/20/2021"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE, warning= FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE)
```

```{r, warning= FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(caret)
library(data.table)
library(knitr)
```

# DO A SIMPLE EDA TO CLEAN THE DATA

```{r, warning=FALSE, message=FALSE}
checkin = read_csv('dailycheckins.csv')
checkin %>%
  head(10) %>%
  kable()
```
Missing values
```{r}
map_dbl(checkin, function(x) sum(is.na(x)))
```

5 missing data from user

### Check user column

```{r}
checkin %>%
        count(user) %>%
  arrange(desc(n)) %>%
  head(15)%>%
  kable()
```

No mistype of entry but with 5 missing data
```{r}
checkin %>%
        filter(is.na(user)) %>%
  kable()
```

Missing users are from different project. The timestamps are identical.

### Check timestamp column

Columns without UTC
```{r}
checkin %>%
        filter(str_detect(timestamp,'UTC', negate = TRUE))
```


```{r}
checkin %>%
        select(timestamp) %>%
        count(str_detect(timestamp,'UTC', negate = TRUE)) %>%
  kable()
```

Can just assume all are UTC

### Count entries with different encoding

```{r}
checkin %>%
        mutate(timestamp = str_remove(timestamp,'UTC')) %>%
        filter(str_detect(timestamp, '[:alpha:]')) %>%
        filter(str_detect(timestamp, 'AM|PM', negate =  TRUE))
```

There are 1022 rows with different encoding.

```{r}
checkin %>%
        mutate(timestamp = str_remove(timestamp,'UTC')) %>%
        filter(str_detect(timestamp, '[:alpha:]')) %>%
        filter(str_detect(timestamp, 'AM|PM', negate =  TRUE)) %>%
        pull(timestamp) %>%
        head(30) %>%
  kable()
```

The words are month in russian.

### Fix timestamp column


Cant detect entries with different encoding as string literal. Manually detect it using filters. Identify those with different encoding
```{r}
checkin = checkin %>%
        mutate(has_utc = str_detect(timestamp, 'UTC')) %>%
        mutate(timestamp = str_remove(timestamp,'UTC')) %>%
        mutate(has_letter = str_detect(timestamp, '[:alpha:]')) %>%
        mutate(notAMPM = str_detect(timestamp, 'AM|PM', negate =  TRUE)) %>%
        mutate(to_change = (has_letter & notAMPM)) %>%
        select(-has_letter, -notAMPM)
```

### Brute force the month since I cant find how to change encoding

Create another table for mapping russian to english
```{r}
checkin = checkin %>%
        mutate(month_russian = str_extract(timestamp, '[:alpha:]+')) %>%
        mutate(month_russian = ifelse(str_detect(month_russian, 'AM|PM'), NA, month_russian))

months = checkin %>%
        distinct(month_russian) %>%
        filter(!is.na(month_russian))


english_months = c('September','November','October','August', 'July','June','May','April','March','February','January','December')

months$english_months = english_months
```

Combine table and change russian characters to english
```{r}
checkin = checkin %>%
        left_join(months, by = c('month_russian' = 'month_russian')) %>%
        mutate(timestamp = ifelse(to_change, str_replace(timestamp,month_russian, english_months), timestamp))
```


Create identifier for those with AMPM and remove AMPM from timestamp
```{r}
checkin = checkin %>%
        mutate(AMPM = str_extract(timestamp, 'AM|PM')) %>%
        mutate(timestamp = ifelse(!is.na(AMPM), str_remove(timestamp, 'AM|PM'), timestamp))
```

Formats
- Those with russian letters is in format Day Month Year Hour Minute. 1022 records
- those with APMP has month, day, year, hour minute format. 1025 records
- those with UTC is in format year month day hour minute second

Create new column fixed timestamp and populate using timestamp
```{r warning=FALSE}
checkin = checkin %>%
        mutate(fixed_timestamp = if_else(has_utc, ymd_hms(timestamp), NA_real_)) %>% #format for utc
        mutate(fixed_timestamp = if_else(to_change, dmy_hm(timestamp), fixed_timestamp)) %>%
        mutate(fixed_timestamp = if_else(!is.na(AMPM), mdy_hm(timestamp), fixed_timestamp))
```

Remove unnecessary columns
```{r}
checkin = checkin %>%
        select(-month_russian, -english_months) %>%
        rename(has_russian = to_change)
```

### Adjust the time based on timezone and AMPM

263 Entries with PM and 762 with AM
```{r}
checkin %>%
        count(AMPM) %>%
  kable()
```
Some AM and PM were just entry errors. Fix timestamp of AM AND PM based on next and previous entries

```{r}
checkin = checkin %>%
        mutate(prev_entry = lead(fixed_timestamp)) %>% #create next and prev timestamp
        mutate(next_entry = lag(fixed_timestamp)) %>%
        mutate(prev_entry_diff = abs(as.duration(fixed_timestamp - prev_entry))) %>% #compute for time difference
        mutate(next_entry_diff = abs(as.duration(next_entry - fixed_timestamp))) %>% 
        mutate(big_diff = prev_entry_diff > hours(3) & next_entry_diff > hours(3)) %>% #counted big difference if both entries are 3 hrs away
        mutate(fixed_timestamp = if_else((big_diff & !is.na(AMPM) & AMPM == 'AM'), fixed_timestamp - hours(12), fixed_timestamp)) %>% #adjust time
        mutate(fixed_timestamp = if_else((big_diff & !is.na(AMPM) & AMPM == 'PM'), fixed_timestamp + hours(12), fixed_timestamp)) 
```


Get only needed columns and adjust time for UTC
```{r}
checkin = checkin %>%
        select(-prev_entry, -next_entry, -next_entry_diff, -prev_entry_diff, -big_diff, -AMPM, -has_russian, -has_utc) %>%
        mutate(fixed_timestamp = fixed_timestamp + hours(8)) %>%
        rename(raw_timesstamp = timestamp) %>%
        rename(timestamp = fixed_timestamp) %>%
        select(timestamp, user, hours, project, raw_timesstamp)
```


### Check hours data

```{r}
summary(checkin$hours)
```

30 hrs?

```{r}
checkin %>%
        filter(hours > 12) %>%
        arrange(desc(hours)) %>%
  head(10) %>%
  kable()
```
All contains the word project. 

## Check project

```{r}
checkin %>%
        count(project) %>%
  arrange(desc(n)) %>%
  head(15) %>%
  kable()
```

There are some typos in the project column

Fix typos
```{r}
checkin = checkin %>%
        mutate(project = str_replace(project, 'blogideas', 'blog-ideas'),
               project = str_replace(project, '^c(.*)and(.*)','cultureandmanagement'),
               project = str_replace(project, 'hirng','hiring'),
               project = str_replace(project, 'internals','internal'),
               project = str_replace(project, 'machine-learning','machinelearning'),
               project = str_replace(project, 'misc','miscellaneous'),
               project = str_replace(project, '^op(.*)min$','opsandadmin'),
               project = str_replace(project, 'pm','projectmanagement'),
               project = str_replace(project, 'workshops','workshop'))
```

Check missing users

```{r}
checkin %>%
        mutate(row = row_number()) %>%
        slice(15790:15810) %>%
  kable()
```
2 missing entries on December 27 with same timein

```{r}
checkin %>%
        mutate(row = row_number()) %>%
        slice(17570:17580) %>%
  kable()
```
Same with last time. Consecutive missing data with same timein. Probably the same person. Cant use previous or next entries for imputation.

There's not much data types to predict 5 missing values


[Onto the EDA](https://github.com/idellang/CastanedaTM_Ans/tree/master/CleaningData)



## Function to summarize data cleaning
```{r}
clean_data = function(data){
        
        data = data %>%
                mutate(has_utc = str_detect(timestamp, 'UTC')) %>%
                mutate(timestamp = str_remove(timestamp,'UTC')) %>%
                mutate(has_letter = str_detect(timestamp, '[:alpha:]')) %>%
                mutate(notAMPM = str_detect(timestamp, 'AM|PM', negate =  TRUE)) %>%
                mutate(to_change = (has_letter & notAMPM)) %>%
                select(-has_letter, -notAMPM)
        
        data = data %>%
                mutate(month_russian = str_extract(timestamp, '[:alpha:]+')) %>%
                mutate(month_russian = ifelse(str_detect(month_russian, 'AM|PM'), NA, month_russian))
        
        months = data %>%
                distinct(month_russian) %>%
                filter(!is.na(month_russian))
        
        
        english_months = c('September','November','October','August', 'July','June','May','April','March','February','January','December')
        
        months$english_months = english_months
        
        data = data %>%
                left_join(months, by = c('month_russian' = 'month_russian')) %>%
                mutate(timestamp = ifelse(to_change, str_replace(timestamp,month_russian, english_months), timestamp))
        
        data = data %>%
                mutate(AMPM = str_extract(timestamp, 'AM|PM')) %>%
                mutate(timestamp = ifelse(!is.na(AMPM), str_remove(timestamp, 'AM|PM'), timestamp))
        
        data = data %>%
                mutate(fixed_timestamp = if_else(has_utc, ymd_hms(timestamp), NA_real_)) %>% #format for utc
                mutate(fixed_timestamp = if_else(to_change, dmy_hm(timestamp), fixed_timestamp)) %>%
                mutate(fixed_timestamp = if_else(!is.na(AMPM), mdy_hm(timestamp), fixed_timestamp))
        
        data = data %>%
                select(-month_russian, -english_months) %>%
                rename(has_russian = to_change)
        
        data = data %>%
                mutate(prev_entry = lead(fixed_timestamp)) %>% #create next and prev timestamp
                mutate(next_entry = lag(fixed_timestamp)) %>%
                mutate(prev_entry_diff = abs(as.duration(fixed_timestamp - prev_entry))) %>% #compute for time difference
                mutate(next_entry_diff = abs(as.duration(next_entry - fixed_timestamp))) %>% 
                mutate(big_diff = prev_entry_diff > hours(3) & next_entry_diff > hours(3)) %>% #counted big difference if both entries are 6 hrs away
                mutate(fixed_timestamp = if_else((big_diff & !is.na(AMPM) & AMPM == 'AM'), fixed_timestamp - hours(12), fixed_timestamp)) %>% #adjust time
                mutate(fixed_timestamp = if_else((big_diff & !is.na(AMPM) & AMPM == 'PM'), fixed_timestamp + hours(12), fixed_timestamp)) 
        
        data = data %>%
                select(-prev_entry, -next_entry, -next_entry_diff, -prev_entry_diff, -big_diff, -AMPM, -has_russian, -has_utc) %>%
                mutate(fixed_timestamp = fixed_timestamp + hours(8)) %>%
                rename(raw_timesstamp = timestamp) %>%
                rename(timestamp = fixed_timestamp) %>%
                select(timestamp, user, hours, project)
        
        data = data %>%
                mutate(project = str_replace(project, 'blogideas', 'blog-ideas'),
                       project = str_replace(project, '^c(.*)and(.*)','cultureandmanagement'),
                       project = str_replace(project, 'hirng','hiring'),
                       project = str_replace(project, 'internals','internal'),
                       project = str_replace(project, 'machine-learning','machinelearning'),
                       project = str_replace(project, 'misc','miscellaneous'),
                       project = str_replace(project, '^op(.*)min$','opsandadmin'),
                       project = str_replace(project, 'pm','projectmanagement'),
                       project = str_replace(project, 'workshops','workshop'))
        
        data
}
```




