---
title: "Exploring Jumping GPS data"
author: "Me"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
```
# Pan

```{r}
jumping_pan = read_csv('output_data/jumping_pan.csv')

jumping_pan %>%
      distinct(plateno)
```
W

```{r}
jumping_pan = jumping_pan %>%
      mutate(speed = distance_covered/(date_part + 1)) %>%
      mutate(kmh = speed * 3.6) %>%
      mutate(invalid_speed = kmh > 100) 

jumping_pan%>%
      group_by(plateno) %>%
      summarise(percent_wrong_data = mean(invalid_speed, na.rm = T)) %>%
      mutate(percent_wrong_data = percent_wrong_data * 100) %>%
      arrange(desc(percent_wrong_data))

jumping_pan %>%
      filter(plateno == 'CAL1825')
```

Only .03 percent have wrong speed
```{r}
jumping_pan %>%
      summarise(mean(invalid_speed, na.rm = T) * 100)
```

There are 283 rows with invalid speed
```{r}
nrow(jumping_pan) * 0.0373459	
```

```{r}
jumping_pan %>%
      filter(date_part > 36000/2)
```

No long datepart 

# EHD

```{r}
jumping_ehd = read_csv('output_data/ehd_jumping.csv')

jumping_ehd = jumping_ehd %>%
      mutate(speed = distance_covered/(date_part + 1)) %>%
      mutate(kmh = speed * 3.6) %>%
      mutate(invalid_speed = kmh > 100) 
```

```{r}
jumping_ehd%>%
      group_by(plateno) %>%
      summarise(percent_wrong_data = mean(invalid_speed, na.rm = T)) %>%
      mutate(percent_wrong_data = percent_wrong_data * 100) %>%
      arrange(desc(percent_wrong_data))
```

```{r}
jumping_ehd %>%
      filter(plateno == 'CAA3526')
```


There are more errorneous data in ehd. There is even a plate number with 97% wrong data.Keep in mind that this data is already processed

```{r}
jumping_ehd %>%
      summarise(mean(invalid_speed, na.rm = T) * 100)
```
32% have invalid speed

Aside from that there were also GPS devices that long date part interval
```{r}
jumping_ehd %>%
      filter(date_part > 36000)
```

# Transpecial


```{r}
jumping_trans = read_csv('output_data/jumping_trans_new.csv')

jumping_trans = jumping_trans %>%
      mutate(speed = distance_covered/(date_part + 1)) %>%
      mutate(kmh = speed * 3.6) %>%
      mutate(invalid_speed = kmh > 100) 
```


```{r}
jumping_trans%>%
      group_by(plateno) %>%
      filter(date_part < 3600) %>%
      summarise(percent_wrong_data = mean(invalid_speed, na.rm = T)) %>%
      mutate(percent_wrong_data = percent_wrong_data * 100) %>%
      arrange(desc(percent_wrong_data))
```
Dahil sa malalayong interval


There are more errorneous data in ehd. There is even a plate number with 97% wrong data.Keep in mind that this data is already processed

```{r}
jumping_trans %>%
      summarise(mean(invalid_speed, na.rm = T) * 100)
```
32% have invalid speed

Aside from that there were also GPS devices that long date part interval
```{r}
jumping_trans %>%
      filter(date_part > 36000)
```

