---
title: "Geofence Verification"
author: "Me"
date: "3/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(RPostgreSQL)
library(getPass)
library(lubridate)
library(data.table)
```


Load the data

```{r}
data = read_csv('input_data/Nestle-CICO-JantoFeb-FIX.csv')
head(data)
```

```{r}
format_data = function(data){
      
      data = data %>%
            select(created, plateno, group_names, alert_msgs,longitude, latitude)
      
      #switch long and lat
      data = data %>%
            mutate(longitude_fix = ifelse(longitude < 100, latitude, longitude),
                   latitude_fix = ifelse(latitude > 100, longitude, latitude)) %>%
            select(-longitude, -latitude) %>%
            rename(longitude = longitude_fix, 
                   latitude = latitude_fix)
      
      return(data)
}

formatted_data = format_data(data)

formatted_data
```


Upload to postgis
```{r}
gps_to_postgis = function(data, postgis_table = 'gps_to_postgis_table', userpass, password){
      pgdrv <- dbDriver(drvName = "PostgreSQL")
      
      postgrescon <-DBI::dbConnect(pgdrv,
                    dbname="WTI",
                    host="localhost", port=5432,
                    user = userpass,
                    password = password)
      
      data = data %>%
            mutate(created = as.character(created))
      
      dbWriteTable(postgrescon, postgis_table, data, row.names=FALSE, append=TRUE)
}

gps_to_postgis(formatted_data, 'geofence_verification_staging', userpass = 'postgres', password = getPass('Password:'))
```

```{r}
formatted_data
```


Query for creating tables in postgresql

CREATE TABLE geofence_verification
(
    created text,
   	geom geometry(point,4326),
	plateno text,
	group_names text,
	alert_msgs text
	
);

INSERT INTO geofence_verification
SELECT 
	created,
	ST_TRANSFORM(ST_SETSRID(ST_POINT(longitude, latitude), 4326), 4326) AS geom,
	plateno,
	group_names,
	alert_msgs
FROM geofence_verification_staging

## Extract data

```{r}
formatted_data %>%
      arrange(plateno, created)
```



```{r}
query = 'SELECT 
	t.created,
	t.plateno,
	t.group_names,
	t.alert_msgs,
	g.name AS geofence_name
FROM 
	geofence_verification_geom AS t
JOIN
	nestle_geofences_updated AS g
ON 
	ST_INTERSECTS(t.geom, ST_BUFFER(g.geom, .0005))
	'

get_gps_inside_geofence = function(userpass, password, query){
   pgdrv <- dbDriver(drvName = "PostgreSQL")
   postgrescon <-DBI::dbConnect(pgdrv,
                    dbname="WTI",
                    host="localhost", port=5432,
                    user = userpass,
                    password = password)
   
   points_in_geofences <- dbGetQuery(postgrescon, query)
 
   points_in_geofences = points_in_geofences %>%
      mutate(created = ymd_hms(created))
      
   points_in_geofences = points_in_geofences %>%
      arrange(created)
     
   return(points_in_geofences)
}

points_inside_geofence = get_gps_inside_geofence(userpass = 'postgres', password = getPass('Password:'), query = query)
```

```{r}
vehicle_location_group = points_inside_geofence %>%
   mutate(track_period = floor_date(created, '1 hour')) %>%
   group_by(plateno, geofence_name, track_period) %>%
   mutate(time_start = min(created),
          time_end = max(created),
          num_points = n()) %>%
   mutate(duration = as.duration(time_end - time_start))
```

```{r}
vehicle_location_group = read_csv('points_grouped.csv')
```

Get distinct groups
```{r}
vehicle_location_group = vehicle_location_group %>%
      select(-created) %>%
      arrange(plateno, geofence_name, track_period) %>%
      distinct()

vehicle_location_group = vehicle_location_group %>%
      rename(dwell_time = duration)
```

```{r}
morethan5mins = vehicle_location_group %>%
         group_by(plateno, geofence_name) %>%
         arrange(plateno, geofence_name)
   
morethan5mins = morethan5mins %>%
   mutate(time_end_previous_location = lag(time_end),
          time_start_next_location = lead(time_start)) %>%
   mutate(time_difference_to_next_location = as.duration(time_start_next_location - time_end),
          time_difference_from_prev_location =as.duration(time_start - time_end_previous_location)) 
```

```{r}
 morethan5mins = morethan5mins %>%
      select(plateno,
             geofence_name,
             time_start, 
             time_end,
             time_end_previous_location,
             time_start_next_location,
             dwell_time,
             time_difference_to_next_location,
             time_difference_from_prev_location,
             everything())
```

```{r}
morethan5mins = morethan5mins %>%
      mutate(continue_next_location = time_difference_to_next_location < 300,
             contine_prev_location = time_difference_from_prev_location < 300)  %>%
    mutate(continue_both_location = contine_prev_location & continue_next_location)

 #filter out data to record only entry and exit 
morethan5mins = morethan5mins %>%
 filter(continue_both_location == F | is.na(continue_both_location))

#compute for actual end time
morethan5mins = morethan5mins %>%
 group_by(plateno, geofence_name) %>%
 mutate(actual_time_end = if_else(
 (is.na(time_end_previous_location) | lead(time_difference_from_prev_location)  < 300) 
 & !(time_difference_to_next_location > 300), 
 lead(time_end),
 time_end
 )) 
#filter no actual time end because they are the last for that location
#select relevant columns
#compute for duration at geofence
morethan5mins = morethan5mins%>%
 filter(!is.na(actual_time_end)) %>%
 select(plateno, geofence_name, time_start, actual_time_end) %>%
 mutate(dwell_time = as.duration(actual_time_end - time_start))

#create ranking based on duration
fixed_data = morethan5mins %>%
 group_by(plateno, geofence_name, actual_time_end) %>%
 mutate(rank = rank(desc(dwell_time)))


#remove the last entry
fixed_data = fixed_data %>%
         filter(rank != 2) 

fixed_data = fixed_data %>%
       ungroup() %>%
       select(plateno, 
              geofence_name,
              first_timestamp = time_start, 
              last_timestamp = actual_time_end, 
              time_duration_at_geofence = dwell_time)
```

```{r}
fixed_data %>%
      arrange(first_timestamp) %>%
      mutate(first_timestamp = as.character(first_timestamp),
             last_timestamp = as.character(last_timestamp)) 
```

```{r}
fixed_data %>%
      write_csv('CICO_OUTPUT.csv')
```

```{r}
fixed_data %>%
      filter(time_duration_at_geofence > 1000)
```


```{r}
geofence_verified = read_csv('CICO_OUTPUT.csv')
geofence_verified %>%
      select(-plateno) %>%
      filter(time_duration_at_geofence > 900) %>%
      distinct(geofence_name) %>%
      write_csv('geofences_at_15min_threshold.csv')
```


```{r}
geofence_verified %>%
      distinct(geofence_name) %>%
      write_csv('All_geofences.csv')
```

