---
title: "Matching Algo cleaning data"
author: "Me"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries
```{r}
library(elastic)
library(elasticsearchr)
library(jsonlite)
library(tidyverse)
library(tidyr)
library(broom)
library(data.table)
library(lubridate)
library(getPass)

conn = connect(
host = "tmsuite-elasticsearch.trackme.com.ph",
port = 443,
path = NULL,
transport_schema = "https",
user = 'jfcastaneda',
pwd =  getPass("Enter Password:"),
headers = NULL,
cainfo = NULL,
force = FALSE,
errors = "simple",
warn = TRUE
)
```

Load cico generated using GPS data
```{r}
data = read_csv('data_sdg_geofence/Transpecial_GPS_CICO_DecToJan.csv')
dim(data)
```

Make a table that looks like FO data
```{r}
data = data %>%
      select(plateno, geofence_name, pattern,
             entry_time = first_timestamp)

cico = data
```

```{r}
cico_to_fo = cico %>%
   select(plateno, geofence_name, entry_time, pattern) %>%
   arrange(plateno, entry_time) %>%
   mutate(source = ifelse(pattern == 1, geofence_name, NA)) %>%
   mutate(source_entrytime = if_else(pattern == 1, as.character(entry_time), NA_character_)) %>%
   group_by(plateno) %>%
   fill(source) %>%
   fill(source_entrytime) %>%
   mutate(consec_pickup = ifelse(((pattern == 1 & lead(pattern) == 1) & plateno == lead(plateno)), TRUE, FALSE))
```

```{r}
cico_to_fo %>%
      write_csv('cico_to_fo.csv')
```


# Process FO data

get arrival and location
```{r}
trans_fo = read_csv('data_sdg_geofence/Transpecial-FO-Jan2021.csv')

trans_fo = trans_fo %>%
   select(trip_number, vehicle_plate_no,so_number,dropoffs, pickups)

trans_fo = trans_fo  %>%
   mutate(dropoffs = str_remove_all(dropoffs,"\\'")) %>%
   mutate(dropoff_date = str_extract(dropoffs,
                                  'arrival: [:digit:]{4}-[:digit:]{2}-[:digit:]{2} [:digit:]{2}:[:digit:]{2}:[:digit:]{2}')) %>%
   mutate(dropoff_name = str_extract(dropoffs,'name: .*')) %>%
   mutate(dropoff_name = str_split(dropoff_name, ',')) %>%
   unnest(dropoff_name) %>%
   filter(str_detect(dropoff_name,'name:'))

trans_fo = trans_fo %>%
     mutate(pickups = str_remove_all(pickups,"\\'")) %>%
   mutate(pickups_date = str_extract(pickups,
                                  'arrival: [:digit:]{4}-[:digit:]{2}-[:digit:]{2} [:digit:]{2}:[:digit:]{2}:[:digit:]{2}')) %>%
   mutate(pickups_name = str_extract(pickups,'name: .*')) %>%
   mutate(pickups_name = str_split(pickups_name, ',')) %>%
   unnest(pickups_name) %>%
   filter(str_detect(pickups_name,'name:')) %>%
   select(-dropoffs, -pickups)

#process dates
trans_fo = trans_fo %>%
   mutate(dropoff_date = ymd_hms(dropoff_date)) %>%
   mutate(pickup_date = ymd_hms(pickups_date)) %>%
   select(-pickups_date)

#clean names
trans_fo = trans_fo %>%
   mutate(dropoff_name = str_remove(dropoff_name,'name:')) %>%
   mutate(pickup_name = str_remove(pickups_name,'name:')) %>%
   select(-pickups_name)
```

# Use data from query

Select needed cols
```{r}
query = "{\"size\": 10000, \"query\": {\"bool\": {\"must\": [], \"filter\": [{\"match_phrase\": {\"client_name\": \"Transpecial Inc.\"}}, {\"nested\": {\"path\": \"pickups\", \"query\": {\"range\": {\"pickups.arrival\": {\"gte\": \"2021-01-01 00:00:00\", \"lt\": \"2021-01-31 23:59:59\", \"time_zone\": \"+08:00\"}}}}}, {\"match_phrase\": {\"group_names\": \"Nestle\"}}, {\"range\": {\"created\": {\"gte\": \"2020-11-30T16:00:00.000Z\", \"lte\": \"2021-02-11T08:17:04.602Z\", \"format\": \"strict_date_optional_time\"}}}], \"should\": [], \"must_not\": []}}}"


data = data = Search(conn, index = 'tmsuite_booking*', body = query, , raw = TRUE)

data = fromJSON(data)
data = data$hits$hits$`_source`

data = data %>%
   select(trip_number, vehicle_plate_no, pickups, dropoffs, so_number)

#get pickup
data = data %>%
   unnest_wider(pickups, names_repair = tidyr_legacy) %>%
   select(trip_number, vehicle_plate_no, so_number,  pickup_arrival = arrival, pickup_name = name, dropoffs)

#get dropoffs
data = data %>%
   unnest_wider(dropoffs, names_repair = tidyr_legacy) %>%
   select(trip_number, vehicle_plate_no, so_number, pickup_arrival, pickup_name, dropoff_arrival = arrival, dropoff_name = name)

transpecial_fo = data
```


# Process bookings data

Get only rows with complete data
```{r}
transpecial_fo = transpecial_fo %>%
   filter(!is.na(pickup_name), !is.na(dropoff_name))
```

Add 8 hours to time
```{r}
transpecial_fo = transpecial_fo %>%
   mutate(pickup_arrival = ymd_hms(pickup_arrival) + hours(8),
          dropoff_arrival = ymd_hms(dropoff_arrival) + hours(8))

transpecial_fo %>%
   summarise(min(pickup_arrival),
             max(pickup_arrival))
# transpecial_fo %>%
#    write_csv('transpecial_fo.csv')
```


# Compare transpecial FO with actual bookings

```{r}
fo_locations = transpecial_fo %>%
   select(pickup_name, dropoff_name) %>%
   stack() %>%
   distinct() %>%
   select(fo_locations = values)

cico_locations = cico %>%
   distinct(geofence_name) %>%
   rename(cico_locations = geofence_name)

cico_locations %>%
   mutate(is_in_fo = .$cico_locations %in% fo_locations$fo_locations) %>%
   filter(is_in_fo == F)
```
There are 42 CICO locations that are not in FO

Try to remove -source destination next time.


Check how many trips were without matching locations
```{r}
cico_to_fo %>%
   select(geofence_name, source) %>%
   filter(!is.na(plateno), !is.na(geofence_name), !is.na(source)) %>%
   mutate(is_in_fo1 = geofence_name %in% fo_locations$fo_locations,
          is_in_fo2 = geofence_name %in% fo_locations$fo_locations) %>%
   filter(is_in_fo1 == T & is_in_fo2 == T) %>%
   filter(geofence_name != source)
```

Only 962 have both their locations in the FO data.


# Process data for exporting
```{r}
cico %>%
   mutate(entry_time = as.character(entry_time)) %>%
   write_csv('trans_cico.csv')

cico_to_fo %>%
   select(source_geofence = source,
          source_entry_time = source_entrytime,
          dest_geofence = geofence_name,
          dest_entry_time = entry_time,
          plateno,
          everything()) %>%
   mutate(source_entry_time = as.character(source_entry_time),
          dest_entry_time = as.character(dest_entry_time)) %>%
   write_csv('trans_cico_fo.csv')

transpecial_fo %>%
   mutate(dropoff_arrival = as.character(dropoff_arrival),
          pickup_arrival = as.character(pickup_arrival)) %>%
   write_csv('trans_fo.csv')
```

















