---
title: "Matching Algo Process"
author: "Me"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data loading

Load libraries
```{r}
library(tidyverse)
library(lubridate)
library(zoo)
library(chron)
```

Load the data
```{r}
trans_cico_fo = read_csv('data_for_matching/trans_cico_to_fo.csv')
trans_cico_fo_ver2 = read_csv('data_for_matching/trans_cico_to_fo_ver2.csv')
trans_fo = read_csv('data_for_matching/trans_fo.csv')
trans_cico = read_csv('data_for_matching/trans_cico.csv')
head(trans_cico_fo)
```


Fix ver 2
```{r}
trans_cico_fo_ver2 = trans_cico_fo_ver2 %>%
   select(source_geofence = source,
          source_entry_time = source_first_timestamp,
          dest_geofence = geofence_name,
          dest_entry_time = first_timestamp,
          plateno,
          everything())
```


segmentized data
```{r}
# remove locations without source
trans_cico_fo = trans_cico_fo_ver2 %>%
      filter(!is.na(source_geofence))

trans_cico_fo = trans_cico_fo_ver2 %>%
      arrange(plateno)

# add segment number if the vehicle returned from trip
trans_cico_fo_segment = trans_cico_fo_ver2 %>%
      filter(pattern == 1) %>%
      mutate(trip_segment = 1:nrow(.))

#left join to add the segment to the trans cico FO
trans_cico_fo_segmentized =  trans_cico_fo_ver2 %>%
      left_join(trans_cico_fo_segment) %>%
      mutate(trip_segment = as.character(trip_segment)) %>%
      mutate(trip_segment = na.locf(.$trip_segment, na.rm = F)) %>%
      filter(source_geofence != dest_geofence)

# add segments
trans_cico_fo_segmentized = trans_cico_fo_segmentized %>%
   mutate(n = 1:nrow(.))
```


Alter trans_cico_segmentized
- add 8 hours
- filter for only January
```{r}
trans_cico_fo_segmentized_8 = trans_cico_fo_segmentized %>%
   mutate(source_entry_time = source_entry_time + hours(8),
          dest_entry_time = dest_entry_time + hours(8))

trans_cico_fo_segmentized_january = trans_cico_fo_segmentized %>%
   filter(source_entry_time >= as.Date('2021-01-01'))
```

get same location and plateno,
if no matching plateno get the closest

```{r}
final_data = tibble()
for (i in 1:819){
  sample = trans_cico_fo_segmentized %>%
   filter(trip_segment == i)

   matched_data = sample %>%
         left_join(trans_fo, by = c('source_geofence' = 'pickup_name', 'dest_geofence' = 'dropoff_name')) %>%
         mutate(same_plateno = vehicle_plate_no == plateno) %>%
         mutate(diff_pickup = abs(as.duration(source_entry_time - pickup_arrival)),
                diff_dropoff = abs(as.duration(dest_entry_time -dropoff_arrival))) %>%
         group_by(source_geofence, dest_geofence) %>%
         mutate(diff_ranking = rank(diff_pickup, ties.method = 'min')) %>%
         arrange(diff_ranking)
   
   
   matched_data = matched_data %>%
      mutate(sum_same_plateno = sum(same_plateno, na.rm = T))
   
   #if no matching plateno, get the closest trip number
   if (all(matched_data$same_plateno == 0, na.rm = T)){
      matched_data = matched_data %>%
         filter(diff_ranking == 1 & !is.na(diff_pickup))
   # if there is matching plateno, get the closest matching plateno
   } else if (all(matched_data$same_plateno == 0, na.rm = T) == F){
      matched_data = matched_data %>%
         filter(diff_ranking == 1 & same_plateno == T & !is.na(diff_pickup))
   }
   
   
   final_data = final_data %>%
               rbind(matched_data)
}

final_data %>%
   select(-pattern, - consec_pickup)
```

There are 380 matches with repeating FOs


Get closest per trip number
- Another approach
- dont FOs after matching but instead after running the algo, find the closest CICO per trip number. 
- Cons: Fewer matches
- Pros : possibly lower time difference
```{r}
final_data %>%
   ungroup() %>%
   filter(!is.na(diff_pickup))
   arrange(trip_number, diff_pickup) %>%
      group_by(source_geofence, dest_geofence, trip_number) %>%
      mutate(pickup_rank = rank(diff_pickup, ties.method = 'min'),
             dropoff_rank = rank(diff_dropoff, ties.method = 'min')) %>%
   filter(pickup_rank == 1 & dropoff_rank == 1) %>%
   arrange(trip_number) %>%
   ungroup() %>%
   summarise(mean_diff_pickup = seconds_to_period(mean(diff_pickup)),
             mean_diff_dropoff = seconds_to_period(mean(diff_dropoff)),
             median_diff_pickup = seconds_to_period(median(diff_pickup)),
             median_diff_dropoff = seconds_to_period(median(diff_dropoff)))
   
```


# Removing matched from the data
```{r}
final_data = tibble()
trans_fo_matching = trans_fo
for (i in 1:819){
  sample = trans_cico_fo_segmentized %>%
   filter(trip_segment == i)

   matched_data = sample %>%
         left_join(trans_fo_matching, by = c('source_geofence' = 'pickup_name', 'dest_geofence' = 'dropoff_name')) %>%
         mutate(same_plateno = vehicle_plate_no == plateno) %>%
         mutate(diff_pickup = abs(as.duration(source_entry_time - pickup_arrival)),
                diff_dropoff = abs(as.duration(dest_entry_time -dropoff_arrival))) %>%
         group_by(source_geofence, dest_geofence) %>%
         mutate(diff_ranking_pickup = rank(diff_pickup, ties.method = 'min'),
                diff_ranking_dropoff = rank(diff_dropoff, ties.method =  'min')) %>%
         arrange(diff_ranking_pickup)
   
   
   matched_data = matched_data %>%
      mutate(sum_same_plateno = sum(same_plateno, na.rm = T))
   
   #if no matching plateno, get the closest trip number
   if (all(matched_data$same_plateno == 0, na.rm = T)){
      matched_data = matched_data %>%
         filter(diff_ranking_pickup == 1 & !is.na(diff_pickup))
   # if there is matching plateno, get the closest matching plateno
   } else if (all(matched_data$same_plateno == 0, na.rm = T) == F){
      matched_data = matched_data %>%
         filter((same_plateno == T) & !is.na(diff_pickup))
   }
   
   
   if (all(matched_data$same_plateno == T, na.rm = T)){
      matched_data = matched_data %>%
         filter(diff_ranking_pickup == min(diff_ranking_pickup),
                diff_ranking_dropoff == min(diff_ranking_dropoff))
   } 
   
   trans_fo_matching =  trans_fo_matching %>%
      anti_join(matched_data, by = c('trip_number' = 'trip_number'))
   
   final_data = final_data %>%
               rbind(matched_data)
}

final_data


# final_data %>%
#    select(-pattern, -consec_pickup) %>%
#    mutate(source_entry_time = as.character(source_entry_time),
#           dest_entry_time = as.character(dest_entry_time),
#           pickup_arrival = as.character(pickup_arrival),
#           dropoff_arrival = as.character(dropoff_arrival)) %>%
#    write_csv('transpecial_removematched.csv')
```



300 matches if no repeating


# Individual matching and removal
```{r}
ptm <- proc.time()

final_data = tibble()
trans_fo_matching = trans_fo
for (i in 1:1379){
  sample = trans_cico_fo_segmentized %>%
   filter(n == 1)

   matched_data = sample %>%
         left_join(trans_fo_matching, by = c('source_geofence' = 'pickup_name', 'dest_geofence' = 'dropoff_name')) %>%
         mutate(same_plateno = vehicle_plate_no == plateno) %>%
         mutate(diff_pickup = abs(as.duration(source_entry_time - pickup_arrival)),
                diff_dropoff = abs(as.duration(dest_entry_time -dropoff_arrival))) %>%
         group_by(source_geofence, dest_geofence) %>%
         mutate(diff_ranking_pickup = rank(diff_pickup, ties.method = 'min'),
                diff_ranking_dropoff = rank(diff_dropoff, ties.method =  'min'))
   
    
   # matched_data = matched_data %>%
   #    mutate(sum_same_plateno = sum(same_plateno, na.rm = T))
   
   #if no matching plateno, get the closest trip number
   if (all(matched_data$same_plateno == 0, na.rm = T)){
      matched_data = matched_data %>%
         filter(diff_ranking_pickup == 1 & !is.na(diff_pickup))
   # if there is matching plateno, get the closest matching plateno
   } else if (all(matched_data$same_plateno == 0, na.rm = T) == F){
      matched_data = matched_data %>%
         filter((same_plateno == T) & !is.na(diff_pickup))
   }
   
   # if there are multiple matching plate no. Get the closest
   # if (nrow(matched_data %>% 
   #          filter(diff_ranking_pickup == min(diff_ranking_pickup))) == 1){
   #    matched_data = matched_data %>%
   #       filter(diff_ranking_pickup == min(diff_ranking_pickup))
   # } 
   # else{
   #    matched_data = matched_data %>%
   #       filter(diff_ranking_pickup == min(diff_ranking_pickup),
   #              diff_ranking_dropoff == min(diff_ranking_dropoff))
   # }
   
   
   # get only the closest
   matched_data = matched_data %>%
      filter(diff_ranking_pickup == min(diff_ranking_pickup))
   
   # if there's a match. Remove those matched with the current FO
   if (nrow(matched_data) > 0){
       trans_fo_matching =  trans_fo_matching %>%
      anti_join(matched_data, by = c('vehicle_plate_no' = 'vehicle_plate_no', 
                                     'pickup_arrival' = 'pickup_arrival', 
                                     'pickup_name' = 'source_geofence',
                                     'dropoff_arrival' = 'dropoff_arrival',
                                     'dropoff_name' = 'dest_geofence'))
   }
   
   
   final_data = final_data %>%
               rbind(matched_data)
}

proc.time() - ptm
final_data
```

## Initial results with improvement on time

```{r}
ptm <- proc.time()

datalist = list()
trans_fo_matching = trans_fo
for (i in 1:1379){
  sample = trans_cico_fo_segmentized_january[trans_cico_fo_segmentized_january$n == i,]

   
   matched_data = left_join(sample, trans_fo_matching, by = c('source_geofence' = 'pickup_name', 'dest_geofence' = 'dropoff_name'))
   
   matched_data$same_plateno = matched_data$vehicle_plate_no == matched_data$plateno
   matched_data$diff_pickup = abs(as.duration(matched_data$source_entry_time - matched_data$pickup_arrival))
   matched_data$diff_dropoff = abs(as.duration(matched_data$dest_entry_time - matched_data$dropoff_arrival))
   
   matched_data = matched_data %>%
         group_by(source_geofence, dest_geofence) %>%
         mutate(diff_ranking_pickup = rank(diff_pickup, ties.method = 'min'),
                diff_ranking_dropoff = rank(diff_dropoff, ties.method =  'min'))
   
   if(all(is.na(matched_data$diff_dropoff))){
      next
   }
   
   # matched_data = matched_data %>%
   #    mutate(sum_same_plateno = sum(same_plateno, na.rm = T))
   
   #if no matching plateno, get the closest trip number
   if (all(matched_data$same_plateno == 0, na.rm = T)){
      matched_data = matched_data[matched_data$diff_ranking_pickup == 1 & !is.na(matched_data$diff_pickup),]
      
   # if there is matching plateno, get the closest matching plateno
   } else if (all(matched_data$same_plateno == 0, na.rm = T) == F){
      matched_data = matched_data[matched_data$same_plateno ==T &!is.na(matched_data$same_plateno),]
   }
   
   # if there are multiple matching plate no. Get the closest
   # if (nrow(matched_data %>%
   #          filter(diff_ranking_pickup == min(diff_ranking_pickup))) == 1){
   #    matched_data = matched_data %>%
   #       filter(diff_ranking_pickup == min(diff_ranking_pickup))
   # }
   # else{
   #    matched_data = matched_data %>%
   #       filter(diff_ranking_pickup == min(diff_ranking_pickup),
   #              diff_ranking_dropoff == min(diff_ranking_dropoff))
   # }
   # 
   
   # get only the closest
   
   matched_data = matched_data[matched_data$diff_ranking_pickup == min(matched_data$diff_ranking_pickup),]
   
   # if there's a match. Remove those matched with the current FO
   if (nrow(matched_data) > 0){
       trans_fo_matching =  trans_fo_matching %>%
      anti_join(matched_data, by = c('vehicle_plate_no' = 'vehicle_plate_no', 
                                     'pickup_arrival' = 'pickup_arrival', 
                                     'pickup_name' = 'source_geofence',
                                     'dropoff_arrival' = 'dropoff_arrival',
                                     'dropoff_name' = 'dest_geofence'))
   }
   
   
  datalist[[i]] = matched_data
}

big_data = do.call(bind_rows, datalist)
proc.time() - ptm
```









