---
title: "Data validation client mapping"
author: "Me"
date: "11/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library(tidyverse)
library(janitor)
```

#load data
```{r}
all_data = read_csv('all_data.csv') %>%
  clean_names() %>%
  select(-x19)
```


#Compare companies and haulers

Get all companies
```{r}
company_cols = all_data %>%
  select(hauling_client_company, conglomerate, company_subsidiary, 
         company, partner_company, client_company, partner_conglomerate, subsidiary_company,
         company_under_industry, customer_company) %>%
  stack() %>%
  distinct(values, ind)%>%
  rename(name = values, column = ind)
```

```{r}
haulers_in_company = all_data %>%
  distinct(customer_hauler) %>%
  mutate(is_in_company = .$customer_hauler %in% company_cols$name) %>%
  filter(is_in_company == T)

haulers_in_company
```

```{r}
company_cols %>%
  filter(name %in% haulers_in_company$hauler_under_industry) %>%
  filter(!is.na(name))
```

Get all haulers


```{r}
haulers_cols = all_data %>%
  select(haulers, customer_hauler, hauler_under_industry, hauler_subsidiary) %>%
  stack() %>%
  rename(name = values, column = ind) %>%
  distinct(name, column)

haulers_cols
```

```{r}
company_cols %>%
  distinct(column)
```


```{r}
company_in_hauler = all_data %>%
  distinct(customer_company) %>%
  filter(!is.na(customer_company)) %>%
  mutate(is_in_hauler = .$customer_company %in% haulers_cols$name) %>%
  filter(is_in_hauler == T)
  
company_in_hauler
```


```{r}
haulers_cols %>%
  filter(name %in% company_in_hauler$company_under_industry) %>%
  filter(!is.na(name))
```

#other misplaced values


```{r}
is_in_company_huler = all_data %>%
  distinct(conglomerate) %>%
  mutate(is_in_company = .$conglomerate %in% company_cols$name) %>%
  mutate(is_in_hauler = .$conglomerate %in% haulers_cols$name) %>%
  filter(is_in_company == T | is_in_hauler == T)

is_in_company_huler
```

```{r}
haulers_cols %>%
  filter(name %in% is_in_company_huler$conglomerate) %>%
  filter(!is.na(name))

company_cols %>%
  filter(name %in% is_in_company_huler$conglomerate) %>%
  filter(!is.na(name))
```



#Check if there are entities with similar names
```{r}
all_data %>%
  select(-units_under_wti)%>%
  stack() %>%
  filter(!is.na(values)) %>%
  distinct(values,ind) %>%
  arrange(values) %>%
  filter(grepl('SM', values))
```

```{r}
all_data %>%
  select(-units_under_wti)%>%
  stack() %>%
  filter(!is.na(values)) %>%
  distinct(values) %>%
  arrange(values)
```
```{r}
all_data %>%
  filter(!is.na(company_connected_to))
```


## no connection company
```{r}
all_data %>%
  select(6:10) %>%
  filter(is.na(partner_company), is.na(partner_conglomerate), is.na(client_company), is.na(subsidiary_company)) %>%
  filter(!is.na(company))
```



#Problematic products

```{r}
all_data %>%
  select(product, customer_company, customer_hauler, status, units_under_wti) %>%
  filter(!is.na(customer_company), !is.na(customer_hauler))
```


#has customer no products
```{r}
all_data %>%
  select(product:units_under_wti) %>%
  filter(!complete.cases(.)) %>%
  filter(is.na(product))
```



#In partner conglomerate not in conglmerate
```{r}
all_data %>%
  select(partner_conglomerate) %>%
  filter(!is.na(.)) %>%
  mutate(is_in_conglo = .$partner_conglomerate %in% all_data$conglomerate)
```



```{r}
all_data %>%
  select(product, customer_hauler, status, units_under_wti) %>%
  distinct(product,customer_hauler) %>%
  arrange(customer_hauler)
```


#customer hauler that has no hauls for
```{r}
all_data %>%
  select(customer_hauler) %>%
  filter(!is.na(customer_hauler)) %>%
  distinct(customer_hauler) %>%
  mutate(has_hauls_for = .$customer_hauler %in% all_data$haulers) %>%
  filter(has_hauls_for == F) %>%
  select(customer_hauler)
```

#FURTHER CHECKING

There are multiple entries in hauler and hauling client company
```{r}
all_data %>%
  distinct(haulers, hauling_client_company) %>%
  group_by(haulers, hauling_client_company) %>%
  count()
```

Products are okay
```{r}
all_data %>%
  distinct(product)
```


There were many haulers with multiple entries
```{r}
all_data %>%
  select(product, customer_hauler) %>%
  group_by(product, customer_hauler) %>%
  count() %>%
  filter(n > 1)
```

```{r}
all_data %>%
  select(conglomerate, company_subsidiary) %>%
  group_by(conglomerate, company_subsidiary) %>%
  count()
```

```{r}
all_data %>%
  select(company, client_company) %>%
  filter(complete.cases(.))
```

```{r}
all_data %>%
  select(industry, company_under_industry) %>%
  group_by(company_under_industry) %>%
  count(industry) %>%
  filter(n > 1)
```


