---
title: "Nestle Trips"
author: "Me"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(RPostgreSQL)
library(getPass)
library(lubridate)
library(data.table)
```

Establish Connection
```{r}

pgdrv <- dbDriver(drvName = "PostgreSQL")
postgrescon <-DBI::dbConnect(pgdrv,
                    dbname="WTI",
                    host="localhost", port=5432,
                    user = 'postgres',
                    password = getPass("Enter Password:"))
```


Process data
```{r}
data = read_csv('GPS-DATA.csv')

data = data %>%
      filter(str_detect(alert_msgs, 'Invalid', negate = TRUE))

data = data %>%
      mutate(datestamp = mdy_hms(datestamp)) %>%
      select(-alert_msgs)
```

Upload data to postgresql
```{r}
dbWriteTable(postgrescon, "nestle_staging", data, row.names=FALSE, append=TRUE)
```

Get data from postgresql

```{r}
points_in_geofences <- dbGetQuery(postgrescon, "SELECT 
	name AS geofence_name,
	datestamp,
	plateno,
	speed
FROM
	nestle_geofence AS g
JOIN
	nestle_gps_data AS t
ON
	ST_INTERSECTS(g.geom, t.geom)	;")

dbDisconnect(postgrescon)
```
```{r}
vehicle_location_group = points_in_geofences %>%
      mutate(track_period = floor_date(datestamp, '30 minutes')) %>%
      group_by(plateno, geofence_name, track_period) %>%
      mutate(time_start = min(datestamp),
             time_end = max(datestamp),
             num_points = n()) %>%
      mutate(duration = as.duration(time_end - time_start)) %>%
      arrange(plateno, geofence_name, track_period) %>%
      filter(num_points > 5) %>%
      select(plateno, geofence_name, track_period, time_start, time_end, duration, num_points)  %>%
      distinct()

vehicle_location_group
```

Check those with low duration and multiple points.

Fivel and plateno 0301603548
```{r}
points_in_geofences %>%
      filter(plateno == '0301603538') %>%
      filter(str_detect(geofence_name, 'Fivel')) %>%
      arrange(datestamp)
```

```{r}
points_in_geofences %>%
      filter(plateno == '0301603538') %>%
      filter(str_detect(geofence_name, 'PH Cagayan')) %>%
      arrange(datestamp) %>%
      filter(datestamp > '2020-12-26 04:00:50')
```


--Hindi doble mabilis lang talaga time interval. Better if icheck yung duration sa loob ng geofence
-- Question: Why are there cases na mas marami yung interval points.

```{r}
#remove those that are not inside the geofence within 5 mins
morethan5mins = vehicle_location_group %>%
      arrange(plateno, track_period) %>%
      filter(duration > duration(minutes = 5))
```
Pwede lagyan ng rank and ifilter yung first per plateno and geofence name

If time end from one time group is less than 2 mins. Make the same time_start
```{r}
#rename column
morethan5mins = morethan5mins %>%
      rename(duration_period_at_geofence = duration)

#add rank
morethan5mins = morethan5mins %>%
      group_by(plateno, geofence_name) %>%
      mutate(rank_place = rank(time_start)) %>%
      arrange(plateno, geofence_name)

#get time difference per group
morethan5mins = morethan5mins %>%
      mutate(time_end_previous = lag(time_end))%>%
      mutate(time_difference_from_previous= as.duration(time_start - time_end_previous)) %>%
      select(-num_points, -rank_place) %>%
      select(plateno, geofence_name, track_period, time_start, time_end, duration_period_at_geofence,time_end_previous, time_difference_from_previous, everything())

#remove those that are staying in the geofence the whole time
morethan5mins = morethan5mins %>%
        mutate(safe = !((duration_period_at_geofence) >= 1500 & time_difference_from_previous <= 300))
```


```{r}


#morethan5mins %>%
     # mutate(time_start_next = lead(time_start),
      #       time_difference_next = as.duration(time_start_next- time_end))


morethan5mins = morethan5mins %>%
      filter(safe == TRUE)




morethan5mins = morethan5mins %>%
      group_by(plateno, geofence_name) %>%
      mutate(actual_time_end = if_else(
            is.na(time_end_previous) | lead(time_difference_from_previous) < duration(minutes = 5), 
            lead(time_end),
            time_end
      )) 


```


```{r}

morethan5mins %>%
      mutate(track_period = as.character(track_period),
             time_start = as.character(time_start),
             time_end = as.character(time_end),
             time_end_previous = as.character(time_end_previous)) %>%
      write_csv('forcheck.csv')

```


```{r}
morethan5mins = morethan5mins%>%
      filter(!is.na(actual_time_end)) %>%
      select(plateno, geofence_name, time_start, actual_time_end) %>%
       mutate(duration_at_geofence = as.duration(actual_time_end - time_start))

fixed_data = morethan5mins %>%
      group_by(plateno, geofence_name, actual_time_end) %>%
      mutate(rank = rank(desc(duration_at_geofence)))

fixed_data = fixed_data %>%
      filter(rank != 2) 

fixed_data = fixed_data %>%
      ungroup() %>%
       mutate(time_start_next = lead(time_start)) %>%
      mutate(time_until_next_location = as.duration(time_start_next - actual_time_end))

fixed_data = fixed_data %>%
      select(plateno, 
             geofence_name,
             first_timestamp = time_start, 
             last_timestamp = actual_time_end, 
             time_duration_at_geofence = duration_at_geofence,
             time_duration_until_next_location = time_until_next_location)
```


Check data in excel
```{r}


fixed_data%>%
      select(- time_duration_until_next_location) %>%
      mutate(first_timestamp = as.character(first_timestamp),
             last_timestamp = as.character(last_timestamp)) %>%
      write_csv('fixed_data.csv')

```


```{r}
morethan5mins = vehicle_location_group %>%
      arrange(plateno, track_period) %>%
      filter(duration > duration(minutes = 5))

morethan5mins = morethan5mins %>%
      rename(duration_period_at_geofence = duration)



#add rank
morethan5mins = morethan5mins %>%
      group_by(plateno, geofence_name) %>%
      mutate(rank_place = rank(time_start)) %>%
      arrange(plateno, geofence_name)

morethan5mins = morethan5mins %>%
      mutate(time_start_next = lead(time_start)) %>%
      mutate(time_difference_next= as.duration(time_start_next - time_end)) %>%
      select(plateno, geofence_name, track_period, time_start, time_end, time_start_next, duration_period_at_geofence, time_difference_next, everything())

morethan5mins %>%
      mutate(track_period = as.character(track_period),
             time_start = as.character(time_start), 
             time_end = as.character(time_end), 
             time_start_next = as.character(time_start_next)) %>%
      write_csv('morethan5mins.csv')
```







