---
title: "Housing Prices"
author: "Me"
date: "11/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load libraries
```{r}
library(tidyverse)
library(caret)
library(ggplot2)
```

#Read data
```{r}
train = read.csv('train.csv')
test = read.csv('test.csv')
```

#Explore data and check for missing values

```{r}
missing_values = map_dbl(train, function(x) sum(is.na(x)))
missing_values = missing_values[missing_values != 0]
missing_values
```
#check missing  value in test

```{r}
map_dbl(test, function(x) sum(is.na(x)))
```
#plot sale price

```{r}
library(ggplot2)
train %>%
  ggplot(aes(SalePrice))+
  geom_histogram()
```
## Check data

Near zero variance columns
```{r}
nz_cols = nearZeroVar(train)
train[,nz_cols]
```

#train processed
```{r}
train_processed = train[,-nz_cols] %>%
  select(-Id) %>%
  mutate(MSZoning = factor(MSZoning)) %>%
  mutate(LotFrontage = coalesce(LotFrontage, 0)) %>%
  mutate(LotArea = as.numeric(LotArea)) %>%
  mutate(Alley = factor(coalesce(Alley, 'None'))) %>%
  mutate(LotShape = factor(LotShape)) %>%
  mutate(LotConfig = factor(LotConfig)) %>%
  mutate(Neighborhood = fct_lump(Neighborhood, 8)) %>%
  mutate(Condition1 = fct_lump(Condition1, 5)) %>%
  mutate(BldgType = factor(BldgType)) %>%
  mutate(HouseStyle = fct_lump(HouseStyle, 5)) %>%
  mutate(OverallQual = factor(OverallQual),
         OverallCond = factor(OverallCond)) %>%
  mutate(RoofStyle = factor(RoofStyle)) %>%
  mutate(Exterior1st = fct_lump(coalesce(Exterior1st,'Other'),6)) %>%
  mutate(Exterior2nd = fct_lump(coalesce(Exterior2nd,'Other'),6)) %>%
  mutate(MasVnrType = factor(coalesce(MasVnrType, 'None'))) %>%
  mutate(MasVnrArea = coalesce(MasVnrArea, 0)) %>%
  mutate(ExterQual = factor(ExterQual),
         ExterCond = factor(ExterCond),
         Foundation = factor(Foundation)) %>%
  mutate(BsmtQual = case_when(
    BsmtQual == 'Ex' ~ 100,
    BsmtQual == 'Gd' ~ 90,
    BsmtQual == 'TA' ~ 80,
    BsmtQual == 'Fa' ~ 70,
    BsmtQual == 'Po' ~ 50,
    TRUE ~ 0
  )) %>%
  mutate(BsmtExposure = factor(coalesce(BsmtExposure,'No'))) %>%
  mutate(BsmtFinType1 = fct_lump(coalesce(BsmtFinType1,'None'),5)) %>%
  mutate(BsmtFinSF1 = coalesce(BsmtFinSF1, 0)) %>%
  mutate(BsmtUnfSF = coalesce(BsmtUnfSF, 0)) %>%
  mutate(TotalBsmtSF = coalesce(TotalBsmtSF, 0)) %>%
  mutate(HeatingQC = factor(HeatingQC))%>%
  mutate(CentralAir = factor(CentralAir)) %>%
  mutate(Electrical = fct_lump(coalesce(Electrical,'Mix'),3)) %>%
  mutate(BsmtFullBath = factor(coalesce(BsmtFullBath,0)))%>%
  mutate(HalfBath = factor(coalesce(HalfBath,0))) %>%
  mutate(BsmtHalfBath = coalesce(BsmtHalfBath,0)) %>%
  mutate(KitchenQual = factor(coalesce(KitchenQual,'Gd'))) %>%
  mutate(FireplaceQu = factor(coalesce(FireplaceQu,'None'))) %>%
  mutate(GarageType = factor(coalesce(GarageType, 'None'))) %>%
  select(-GarageYrBlt) %>%
  mutate(GarageFinish = factor(coalesce(GarageFinish,'None'))) %>%
  mutate(PavedDrive = factor(PavedDrive))%>%
  mutate(PoolQC = fct_lump(coalesce(PoolQC,'None'),2)) %>%
  mutate(Fence = factor(coalesce(Fence, 'None'))) %>%
   mutate(MoSold = factor(MoSold),
         SaleCondition = factor(SaleCondition)) %>%
  mutate(SaleType = fct_lump(coalesce(SaleType,'Oth'),3)) %>%
  mutate(GarageArea = coalesce(GarageArea, 0)) %>%
  mutate(GarageCars = coalesce(GarageCars,0))
  
```



#Convert test data
```{r}
test_processed = test[-nz_cols] %>%
  select(-Id) %>%
  mutate(MSZoning = factor(coalesce(MSZoning, 'RL'))) %>%
  mutate(LotFrontage = coalesce(LotFrontage, 0)) %>%
  mutate(LotArea = as.numeric(LotArea)) %>%
  mutate(Alley = factor(coalesce(Alley, 'None'))) %>%
  mutate(LotShape = factor(LotShape)) %>%
  mutate(LotConfig = factor(LotConfig)) %>%
  mutate(Neighborhood = fct_lump(Neighborhood,8)) %>%
  mutate(Condition1 = fct_lump(Condition1, 5)) %>%
  mutate(BldgType = factor(BldgType)) %>%
  mutate(HouseStyle = fct_lump(HouseStyle, 5)) %>%
  mutate(OverallQual = factor(OverallQual),
         OverallCond = factor(OverallCond)) %>%
  mutate(RoofStyle = factor(RoofStyle)) %>%
  mutate(Exterior1st = fct_lump(coalesce(Exterior1st,'Other'),6)) %>%
  mutate(Exterior2nd = fct_lump(coalesce(Exterior2nd,'Other'),6)) %>%
  mutate(MasVnrType = factor(coalesce(MasVnrType, 'None'))) %>%
  mutate(MasVnrArea = coalesce(MasVnrArea, 0)) %>%
  mutate(ExterQual = factor(ExterQual),
         ExterCond = factor(ExterCond),
         Foundation = factor(Foundation)) %>%
  mutate(BsmtQual = case_when(
    BsmtQual == 'Ex' ~ 100,
    BsmtQual == 'Gd' ~ 90,
    BsmtQual == 'TA' ~ 80,
    BsmtQual == 'Fa' ~ 70,
    BsmtQual == 'Po' ~ 50,
    TRUE ~ 0
  )) %>%
  mutate(BsmtExposure = factor(coalesce(BsmtExposure,'No'))) %>%
  mutate(BsmtFinType1 = fct_lump(coalesce(BsmtFinType1,'None'),5)) %>%
  mutate(BsmtFinSF1 = coalesce(BsmtFinSF1, 0)) %>%
  mutate(BsmtUnfSF = coalesce(BsmtUnfSF, 0)) %>%
  mutate(TotalBsmtSF = coalesce(TotalBsmtSF, 0)) %>%
  mutate(HeatingQC = factor(HeatingQC)) %>%
  mutate(CentralAir = factor(CentralAir)) %>%
  mutate(Electrical = fct_lump(coalesce(Electrical,'Mix'),3)) %>%
  mutate(BsmtFullBath = factor(coalesce(BsmtFullBath,0))) %>%
  mutate(HalfBath = factor(coalesce(HalfBath,0))) %>%
  mutate(BsmtHalfBath = coalesce(BsmtHalfBath,0)) %>%
  mutate(KitchenQual = factor(coalesce(KitchenQual,'Gd'))) %>%
  mutate(FireplaceQu = factor(coalesce(FireplaceQu,'None'))) %>%
  mutate(GarageType = factor(coalesce(GarageType, 'None'))) %>%
  select(-GarageYrBlt) %>%
  mutate(GarageFinish = factor(coalesce(GarageFinish,'None'))) %>%
  mutate(PavedDrive = factor(PavedDrive)) %>%
  mutate(PoolQC = fct_lump(coalesce(PoolQC,'None'),2)) %>%
  mutate(Fence = factor(coalesce(Fence, 'None'))) %>%
   mutate(MoSold = factor(MoSold),
         SaleCondition = factor(SaleCondition)) %>%
  mutate(SaleType = fct_lump(coalesce(SaleType,'Oth'),3)) %>%
  mutate(GarageArea = coalesce(GarageArea, 0))  %>%
  mutate(GarageCars = coalesce(GarageCars,0))

```

 
#check if there are other missing
```{r}
test_missing = map_dbl(test_processed, function(x) sum(is.na(x)))
train_missing = map_dbl(train_processed, function(x) sum(is.na(x)))
train_missing[train_missing !=0]
print('---')
test_missing[test_missing !=0]

```
# Further_processing
```{r}
train_x = train_processed %>%
  select(-SalePrice)

train_y = train_processed$SalePrice
```


# Remove high correlation in numeric
```{r}
nz_cols = nearZeroVar(train_x)
train_x = train_x[,-nz_cols]
test_processed = test_processed[,-nz_cols]
```


```{r}
nums <- unlist(lapply(train_x, is.numeric)) 
names_nums = nums[nums == T] %>%
  names()

numeric_train = train_x[,names_nums]
high_cor = findCorrelation(cor(numeric_train), cutoff = .7)
high_cor = names(numeric_train)[high_cor]
train_x = train_x %>%
  select(-high_cor)
test_processed = test_processed %>%
  select(-high_cor)
```
#parallel computing

```{r}
library(doParallel)
cl = makeCluster(8)
registerDoParallel(cl)

```

#train gbm model
```{r}
gbmGrid <- expand.grid(interaction.depth = seq(1, 7, by = 2),
                       n.trees = seq(100, 1000, by = 50),
                       shrinkage = c(0.01, 0.1),
                       n.minobsinnode = 10)
set.seed(100)
gbmTune <- train(x = train_x, y = train_y,
                 method = "gbm",
                 tuneGrid = gbmGrid,
                 verbose = FALSE)
gbmTune
```

#check RMSE
```{r}
RMSE(predict(gbmTune), train_y)
```

```{r}
preds = predict(gbmTune, test_processed)
submission %>%
  mutate(SalePrice = preds) %>%
  write_csv('my_submission.csv')
```


```{r}
numeric_train = map_lgl(train_x, is.numeric)
numeric_names = numeric_train[numeric_train == T] %>%
  names()

numeric_train = train_x[,numeric_names]
```

# use another method
```{r}
library(e1071)
map_dbl(numeric_train, skewness)
```


#transform test
```{r}
trans = preProcess(train_x, method = c('BoxCox','center','scale'))
transformed_train_x = predict(trans, train_x)
```

#transform train
```{r}
trans = preProcess(test_processed, method = c('BoxCox','center','scale'))
transformed_test= predict(trans, test_processed)
```

#gbm mode
```{r}
gbmGrid <- expand.grid(interaction.depth = seq(1, 7, by = 2),
                       n.trees = seq(100, 1000, by = 50),
                       shrinkage = c(0.01, 0.1),
                       n.minobsinnode = 10)
set.seed(100)
gbmTune <- train(x = transformed_train_x, y = train_y,
                 method = "gbm",
                 tuneGrid = gbmGrid,
                 verbose = FALSE)
gbmTune
```
```{r}
preds = predict(gbmTune, transformed_test)
submission = read_csv('sample_submission.csv')

submission %>%
  mutate(SalePrice = preds) %>%
  write_csv('gbm_trans.csv')
```

#try random forest

#convert to character
```{r}
train_x = data.frame(lapply(train_x, as.character), stringsAsFactors=FALSE)
train_y
```

#convert to character
```{r}
test_processed = data.frame(lapply(test_processed, as.character), stringsAsFactors=FALSE)
```


#create random forest tune
```{r}
mtryGrid <- data.frame(mtry = floor(seq(10, ncol(train_x), length = 8)))

set.seed(100)
rfTune <- train(x = train_x, y = train_y,
                method = "rf",
                tuneGrid = mtryGrid,
                ntree = 500)
rfTune

plot(rfTune)
```

#write to csv
```{r}
preds = predict(rfTune, test_processed)
submission = read_csv('sample_submission.csv')
submission %>%
  mutate(SalePrice = preds) %>%
  write_csv('rf.csv')
```










